The effects of binarity on planet occurrence rates measured by transit surveys

### 2017-10-27 paper/figures.tex
wordsmithed text

\begin{figure}
    \begin{center}
        \includegraphics[width=\textwidth]{figures/errcases_rate_density_vs_radius_model_1_brokenx.pdf}
    \end{center}
    \caption{
    Inferred planet occurrence rates as a function of planet radius in Model 
    \#1.
    This model has fixed stars, fixed planets, twin binaries;
    Note the break in the $x$ axis.
    If the true planet radius is $r_p$, all planets 
    detected in binaries will have apparent radii $r_a = r_p/\sqrt{2}$.
    Only 1 in 8 selected binaries is actually searchable (see 
    Sec.~\ref{sec:model_1}).
    }
    \label{fig:errcases_model_1}
\end{figure}

@@

>>>>

\newpage
\begin{figure}
    \begin{center}
        \includegraphics[width=\textwidth]{figures/errcases_rate_density_vs_radius_model_1_brokenx.pdf}
    \end{center}
    \caption{
    Inferred planet occurrence rates as a function of planet radius in Model 
    \#1.
    This model has fixed stars, fixed planets, and twin binaries.
    If the true planet radius is $r_p$, all planets 
    detected in binaries will have apparent radii $r_a = r_p/\sqrt{2}$.
    Only 1 in 8 selected binaries is actually searchable (see 
    Sec.~\ref{sec:model_1}).
    To help illustrate the individual effects of errors, we 
    separated them:
    ``only wrong $N_\star$'' means the only error is an incorrectly assumed 
    number of selected stars;
    ``only wrong $Q$'' means the only error is an incorrectly assumed 
    detection efficiency (including both miscalculated transit probabilities 
    and 
    fraction of selected stars that are searchable);
    ``only wrong $r_a$'' means the only error is misinterpretation of 
    planetary radii, due to both transit depth dilution and also wrongly 
    assumed host star radii.
    }
    \label{fig:errcases_model_1}
\end{figure}

@@

\begin{figure}[!tb]
    \centering
    \includegraphics[width=.9\textwidth]{figures/HJ_correction_inputrate_vs_HJratevalues.pdf}
    \caption{
    $\Lambda_i$ is the occurrence rate integrated over all possible phase 
    space for the $i^{\rm th}$ system type. $\Lambda_2/\Lambda_0=1$ 
    corresponds to an equal number of planets per secondary as per single star;
    $\Lambda_2/\Lambda_0=0$ corresponds to secondaries not having any planets.
    In our Model \#3, though the true HJ occurrence rate 
    (Eq.~\ref{eq:occ_rate}) is highly dependent on $\Lambda_2$, 
    the inferred rate hardly depends on whether secondaries have HJs.
    This means that the ``correction factor'' between the inferred rate and 
    the true rate around single stars is underestimated by a factor of 
    $\sim1.3$, independent of the HJ rate around secondaries.
    }
    \label{fig:HJ_correction_inputrate_vs_HJratevalues}
\end{figure}

### 2017-10-27 paper/model_2.tex
wordsmithed text

The rate density in this model, $\Gamma(\vec{x})$, is the sum over system 
types of $w_i \Lambda_i p_i(\vec{x})$:
\begin{equation}
\Gamma(\vec{x})
=
\delta^4(r_p,R_\star,a_p,P_p)(w_0 \Lambda_0 + w_1 \Lambda_1)
+ w_2 \Lambda_2 \delta^3(r_p, P_p, a_p) p_2(q),
\label{eq:model2_rate_density}
\end{equation}
where the semimajor axis of the planet must be such that its period is $P_p$, 
and $p_2(q)$ is expressed in terms of the mass ratio instead of the secondary 
star's radius for convenience ($q$ and $R_2$ are interchangeable).
The probability that a secondary hosts a planet, as a function of the mass 
ratio, is
\begin{align}
p_2(q) &= p({\rm has\ planet}\,|\,{\rm secondary\ with\ }q) \times
          p({\rm secondary\ with\ }q)
          \\
p_2(q) &\propto q^{\gamma + \beta} (1+q^\alpha)^{3/2}.
\label{eq:model_2_p_2}
\end{align}
We take first term, $p({\rm has\ planet}\,|\,{\rm secondary\ with\ }q)$, as a 
power law of $q$ with exponent $\gamma$.
For the second term, since the selected sample at a given $(r,P,a)$ is
magnitude-limited, $p({\rm secondary\ with\ }q)$ 
is the product of the volume limited probability and a 
Malmquist-like bias $(1+q^\alpha)^{3/2}$.
In passing, we note that various authors (ourselves included) have incorrectly 
used volume-limited binary distributions in Monte Carlo simulations of 
transit surveys (Sullivan et al 2015, Hartman and Bakos XX, Guenther et al 
2017, Bouma et al 2017).

>>>>

The rate density in this model, $\Gamma(\vec{x})$, is the sum over system 
types of $w_i \Lambda_i p_i(\vec{x})$:
\begin{equation}
\Gamma(\vec{x})
=
\delta^4(r_p,R_\star,a_p,P_p)(w_0 \Lambda_0 + w_1 \Lambda_1)
+ w_2 \Lambda_2 \delta^3(r_p, P_p, a_p) p_2(q),
\label{eq:model2_rate_density}
\end{equation}
where the semimajor axis of the planet must be such that its period is $P_p$, 
and $p_2(q)$ is expressed in terms of the mass ratio instead of the secondary 
star's radius for convenience ($q$ and $R_2$ are interchangeable).
The probability that a secondary hosts a planet, as a function of the mass 
ratio, is
\begin{align}
p_2(q) &= p({\rm has\ planet}\,|\,{\rm secondary\ with\ }q) \times
          p({\rm secondary\ with\ }q)
          \\
p_2(q) &\propto q^{\gamma + \beta} (1+q^\alpha)^{3/2}.
\label{eq:model_2_p_2}
\end{align}
We take first term, $p({\rm has\ planet}\,|\,{\rm secondary\ with\ }q)$, as a 
power law of $q$ with exponent $\gamma$.
For the second term, since the selected sample at a given $(r,P,a)$ is
magnitude-limited, $p({\rm secondary\ with\ }q)$ 
is the product of the volume limited probability, $q^\beta$, and a 
Malmquist-like bias, $(1+q^\alpha)^{3/2}$.
We note that various authors (ourselves included) have incorrectly 
used volume-limited binary distributions in Monte Carlo simulations of 
transit surveys (Sullivan et al 2015, Hartman and Bakos XX, Guenther et al 
2017, Bouma et al 2017).

### 2017-10-27 paper/model_3.tex
wordsmithed text

In this model, as in the previous one, all single and primary stars have 
identical properties.
Only the secondaries have masses, radii, and luminosities that vary between 
systems: $M\propto R \propto L^{1/\alpha}$.
The radii of planets are assigned independently of any host system
property, and are sampled from an intrinsic radius distribution, which we take 
as
\begin{align}
p_r(r)
&\propto
\left.
\begin{cases}
r^\delta & \text{for } r\geq 2R_\oplus \\
{\rm constant} & \text{for } r\leq2R_\oplus.
\end{cases}
\right.
\label{eq:model3_radius_distribution}
\end{align}
Following Howard et al. (2012)'s measurement, we take $\delta = -2.92$.
Our ``nominal model'' remains the same: the binary fraction is 0.44, 
$\alpha=\beta=\gamma=0$.
We take $\Lambda_i$, the occurrence rate integrated over all possible phase 
space for the $i^{\rm th}$ system type, to be equal for singles, primaries, 
and secondaries.

For this model, we forgo analytic development and focus only on numerics 
(Sec.~\ref{sec:numerical_methods}).
The nominal results are shown in Fig.~\ref{fig:errcases_model_3_log}.
For the assumed planetary and stellar distributions, the inferred rate is 
underestimated over all radii.
Taking Fig.~\ref{fig:errcases_model_3_log} and counting the number of planets 
per star with $r>8R_\oplus$, we can compare the true and inferred hot Jupiter 
occurrence rates.
Under the previously described assumptions, the true rate is 9.1 hot Jupiters 
per thousand stars.
The inferred rate is 6.9 per thousand stars.
This means that the inferred rate underestimates the true rate by a 
multiplicative factor of $\sim 1.3$.
This number is suggestively close to the claimed discrepancy between 
hot Jupiter occurrence rates measured by radial velocity and transit surveys 
(Wright et al. 2012).

However, this result only applies under the assumption that $\Lambda_0 = 
\Lambda_1 = \Lambda_2$.
Given that hot Jupiters are thought to be less common around less-massive 
stars, it would be more sensible to consider $\Lambda_2<\Lambda_0$, while 
letting single stars and primaries host planets at the same rate.
We show the HJ multiplicative correction factor as a function of 
$\Lambda_2/\Lambda_0$ in Fig BLAH.

>>>>

In this model, as in the previous one, all single and primary stars have 
identical properties.
Only the secondaries have masses, radii, and luminosities that vary between 
systems: $M\propto R \propto L^{1/\alpha}$.
The radii of planets are assigned independently of any host system
property, and are sampled from an intrinsic radius distribution, which we take 
as
\begin{align}
p_r(r)
&\propto
\left.
\begin{cases}
r^\delta & \text{for } r\geq 2r_\oplus \\
{\rm constant} & \text{for } r\leq2r_\oplus.
\end{cases}
\right.
\label{eq:model3_radius_distribution}
\end{align}
Following Howard et al. (2012)'s measurement, we take $\delta = -2.92$.
Our ``nominal model'' remains the same: the binary fraction is 0.44, 
$\alpha=\beta=\gamma=0$.
We take $\Lambda_i$, the occurrence rate integrated over all possible phase 
space for the $i^{\rm th}$ system type, to be equal for singles, primaries, 
and secondaries.

For this model, we forgo analytic development and focus only on numerics 
(Sec.~\ref{sec:numerical_methods}).
The nominal results are shown in Fig.~\ref{fig:errcases_model_3_log}.
For the assumed planetary and stellar distributions, the inferred rate is 
underestimated over all radii.
Taking Fig.~\ref{fig:errcases_model_3_log} and counting the number of planets 
per star with $r>8r_\oplus$, we can compare the true and inferred hot Jupiter 
occurrence rates.
Under the previously described assumptions, the true rate is 9.1 hot Jupiters 
per thousand stars.
The inferred rate is 6.9 per thousand stars.
This means that the inferred rate underestimates the true rate by a 
multiplicative factor of $\sim 1.3$.
This number is suggestively close to the claimed discrepancy between 
hot Jupiter occurrence rates measured by radial velocity and transit surveys 
(Wright et al. 2012).

However, this result only applies under the assumption that $\Lambda_0 = 
\Lambda_1 = \Lambda_2$.
If hot Jupiters are less common around lower mass stars, it would be more 
sensible to consider $\Lambda_2<\Lambda_0$, while letting single stars and 
primaries host planets at the same rate.
Therefore in Fig.~\ref{fig:HJ_correction_inputrate_vs_HJratevalues}, we let 
$\Lambda_2$ vary, and show the resulting inferred and true hot Jupiter 
($r>8r_\oplus$) rates.
We see that the inferred rate is nearly independent of $\Lambda_2$~--~this is 
because most ($<1/10$) secondaries are not searchable, and so their 
completeness fraction is much smaller than that of primaries or single stars.
This means far fewer detected hot Jupiters orbit secondaries, and so they 
hardly affect the inferred rate.
While the true rate across the entire population is highly dependent on 
$\Lambda_2$, as one would expect from Eq.~\ref{eq:occ_rate}, the rate around 
singles and primaries (green line) is independent of that around secondaries.

This means that, in this case, the inferred rate should be revised upward by a 
factor of $\sim1.3$, independent of the HJ rate around secondaries.
Assuming that RV surveys are measuring the true rate around single stars (or 
primaries), this suggests that the effect appreciably contributes to the 
HJ rate discrepancy.
